### 2018-03-13 experiment_1/experiment_1.tex
new results

\documentclass[11pt,letterpaper]{article}
\usepackage{fullpage}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{float}
\usepackage{amsfonts}
\usepackage{algorithm,algpseudocode}
\usepackage{pdfpages}

\graphicspath{{../results}}

\begin{document}

\title{Experiment 1}
\maketitle

\section*{Simulation Details}

Considered $K = 3$, $T = 1001$, $N = 1200$. Report statistics at $t = 1000$ \\
\textbf{The Bandit priors that were considered}:
\begin{itemize}
\item Uniform: Draw the mean rewards for the arms from [0.25, 0.75]
\item ``HeavyTail": We took the mean rewards to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$). With this distribution it was likely to have arms that were at the extremes (close to 1 and close to 0) but also some of the arms with intermediate value means.
\item Needle-in-haystack
\begin{enumerate}
\item High - 2 arms with mean 0.50, 1 arm with mean 0.70 (+ 0.20)
\end{enumerate}
\end{itemize}
\textbf{Algorithms considered}:
\begin{enumerate}
\item ThompsonSampling with priors of $Beta(1, 1)$ for every arm.
\item DynamicGreedy with priors of $Beta(1, 1)$ for every arm
\item Bayesian Dynamic $\epsilon$-greedy with priors of $Beta(1, 1)$ for every arm and $\epsilon=0.05$
\end{enumerate}
\textbf{Agent Algorithms considered}:
\begin{enumerate}
\item HardMax
\item HardMaxWithRandom
\item SoftMax ($\alpha = 30$)
\end{enumerate}
\textbf{Memory Sizes}
\begin{enumerate}
\item 100
\end{enumerate}
\pagebreak
\textbf{Simulation Procedure}
\begin{algorithm}
\begin{algorithmic}[1]
\For{Each prior $p$}
\State Generate true distribution from $p$ (except for needle-in-haystack, just use $p$ itself)
\State Generate $T \times K$ realizations for the arms 
	\For{Each agent algorithm $agent alg$}
		\For{Each principal algorithm pair $principalalg1$, $principalalg2$}
			\For{$N$ simulations}

				\State Give the agents 5 observations from each principal
				\State Run simulation for T periods
			\EndFor
		\EndFor
	\EndFor
\EndFor
\end{algorithmic}
\end{algorithm}

\section*{Results}
All results are reported for memory size = 100. \\
\vspace{0.25cm}

The rows represent the agent algorithms (defined as above). The columns represent principal 1 vs principal 2 (e.g. TS vs DG means Thompson Sampling is principal 1 and Dynamic Greedy is principal 2).  \\
\vspace{0.1cm}

Within each cell, the following data are presented: \\

\begin{enumerate}
\item The first row displays the sample mean market share for principal 1 as well as a 95\% confidence band of the mean.

\item The second row displays the sample variance of the market share and in parentheses are 95 \% confidence bands for the variance.

\item The third row displays the \% of simulations that resulted in ``extreme" market shares. Extreme market shares are defined as being simulations were one of the principals (either principal 1 or 2) ended up with 90\% or more of the market.
\end{enumerate}
\includepdf[pages={-}]{agents_table}
Comments: Almost everything is just a 50-50 split!
\begin{enumerate}
\item Almost all of the same algorithms pitted against each other ended with getting a 50/50 market share. This seems to be a good sanity check.
\item Strange things: focus on TS vs DEG (Heavy Tail). DEG wins for HM and HMR, but TS wins for SM. This seems weird? TS also seems to lose on HMR with Uniform prior.
\item Why does DynamicGreedy do so well on the HeavyTail prior? Does this make sense?
\end{enumerate}

\end{document}