### 2018-06-07 new_overview/new_experiments.tex
more experiments

\documentclass[11pt,letterpaper]{article}
\usepackage{fullpage}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{float}
\usepackage{amsfonts}
\usepackage{algorithm,algpseudocode}
\usepackage{pdfpages}

\graphicspath{{../results}}

\begin{document}

\title{New Experiments}
\maketitle

\section*{.5/.7 Instance}

plots plots

\section*{Vary Warm Start + Effective end of game}

Here we vary
\includepdf[pages={-}]{eeog_old_priors}

\section*{Various Needles in a Haystack}

\includepdf[pages={-}]{haystack_eeog}

\end{document}