### 2018-07-31 new_conjectures/results_overview.tex

\documentclass[11pt,letterpaper]{article}
\usepackage{fullpage}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{float}
\usepackage{amsfonts}
\usepackage{algorithm,algpseudocode}
\usepackage{pdfpages}

\graphicspath{{../results}}

\begin{document}

\title{Competing Bandits - Summary}
\maketitle

\section*{Overview of Results}

\textbf{Priors}:
\begin{enumerate}
\item Needle In a Haystack (x, K) = (K-1) arms with a mean of x and 1 arm with mean x + 0.2
\item Heavy Tail - means of arms randomly drawn from beta(0.6, 0.6)
\item Uniform - means of arms randomly drawn from [0.25, 0.75]
\item .5/.7 Prior - means of arms randomly drawn from $\{0.5, 0.7\}$
\end{enumerate}

\textbf{Question}: What algorithms win when both firms start at the same time (simultaneous entry)? How does this vary across agent response functions?

\begin{enumerate}
\item Can we use any metrics of the bandit instances run in isolation that help us predict who wins in the competion game? A metric that seems to work well is the relative reputation plot that simply computes the proportion of instances when an algorithm $a$ has at least as high of a reputation than algorithm $b$ at fixed time step $t$ and on the same bandit instance (with the same realizations).
\begin{enumerate}
\item For HM - fixing a warm start value $t$ (number of free observations at the start), whichever algorithm has $>0.5$ proportion of ``victories" at time $t$ according to the relative reputation plot should, on average, win more in competition
\item For HMR and SM, whichever algorithm has $>0.5$ proportion of ``victories" for sufficiently large $t$ (i.e. when learning is more or less done) should win in competition. This is the case since the reputation curve of a given algorithm eventually converges and we simply need to ask which reputation curve is relatively higher at the convergence value.
\item \textbf{Question}: Does this always line up with looking at the mean reputation plots? In one case, Heavy Tail with $K = 3$, we see that the mean plots show $DEG > DG$, but our simulation results show that $DG > DEG$ in competition and $DG > DEG$ in the relative reputation plots. Relation to Dueling Algorithms paper (?). One intuition to this which plays out in the data is that DG either:
\begin{enumerate}
\item identifies the best arm and thus beats or equals reputation of DEG since DEG explores randomly with $\epsilon$ probability.
\item does not identify the best arm and does substantially worse than DEG.
\end{enumerate}
As a result, it will have a worse average reputation due to the cases where it identifies the wrong best arm but there may be enough cases where it does so that when only comparing relative performance in a given simulation it does better.
\end{enumerate}
\item What algorithms win when fixing the HM response model?
\begin{enumerate}
\item We see that there is a low effective end of game\footnote{Effective End of Game (EEOG) is defined as the last round when the agent ``switched" the firm she chose}, which is skewed with many instances having a median of 0, meaning that many of the games end simply from the choices made in the warm start
\item Many extreme shares - one principal takes most of the market
\item What algorithm wins depends on the warm start parameter (number of initial free observations). 
\item For low warm start of 5 we have the following results
\begin{enumerate}
\item $TS > DEG, DG$ for Needle In Haystack (0.5, 10), (0.7, 10)
\item 50/50 for Needle In Haystack (0.5, 3), (0.1, 10), (0.3, 10),  Uniform (K = 3, 10), .5/.7 prior
\item $DEG, DG > TS$ for Heavy Tail (K = 3, 10)
\item $DG > DEG$ for Heavy Tail (K = 3)
\end{enumerate}
\item For moderate warm start (20, 50)
\begin{enumerate}
\item $DEG, DG > TS$ for Uniform (K = 10), Heavy Tail (K = 10)
\item $TS > DEG, DG$ for Needle In Haystack (0.5, 10), (0.7, 10)
\end{enumerate}
\item For large warm start
\begin{enumerate}
\item $TS > DEG, DG$ for everything except .5/.7 prior (why not for .5/.7 prior? Will run more simulations to see if it narrowly wins)
\end{enumerate}
\item \textbf{Question}: Why do results vary by warm start value?
\begin{enumerate}
\item On ``harder" learning instances, learning takes longer and pure exploration has early reputation costs. Warm start needs to be sufficiently high to have recovered the reputation costs of exploration.
\item On ``harder" learning instances, there is a benefit to exploration in the sense that it allows the principal to differentiate itself in reputation against an algorithm like dynamic greedy in the long run.
\item On ``easier" learning instances, little reputation cost but also little benefit to deploying a smarter learning algorithm?
\item How to define ``harder?" Use  $H = \mathbb{E}[\sum\limits_{\mu_{i} < \mu^{*}} \frac{1}{\mu^{*} - \mu_{i}}]$, where $\mu^{*}$ is the mean of the best arm to define instance difficulty. On the ``harder" instances we seem to see larger warm start values required in order for the better algorithm to win but not sure this explains everything.
\end{enumerate}
\end{enumerate}

\item For HMR:
\begin{enumerate}
\item ``Better algorithm" wins with sufficiently large time horizon, where ``better algorithm" is defined as the algorithm that has a larger proportion of wins in the relative reputation plot
\item One exception is the .5/.7 prior (where means of the arms are randomly drawn from {0.5, 0.7}) where even this is 50/50
\item Moderate variance in the market shares
\end{enumerate}
\item For SM

\begin{enumerate}
\item Qualitatively similar to HMR where better algorithm wins for sufficiently large time horizon
\item Results closer to 50/50 than HMR
\item Very low variance in the shares
\end{enumerate}

\end{enumerate}

\textbf{Question}: What happens if there is asymmetric entry so that one principal has an incumbency advantage?

\begin{enumerate}
\item If the incumbency advantage is sufficiently long, then TS is a dominant strategy across priors and agent models that we have tested out.
\item If the incumbency advantage is not sufficiently long, TS is no longer a dominant strategy
\item For HM, TS leads to one firm (the incumbent) dominating the market.
\item For HMR, TS leads to one firm dominating the market but not as dramatically compared to HM.
\item For SM, TS still wins but closer to 50/50. than HMR/HM.
\item For the entrant, it is ambiguous what is the best strategy
\end{enumerate}

\textbf{Question}: Incumbency gives both an informational and a reputational advantage. If we artificially erase one upon entry of the entrant, does information or reputation play a bigger role as barriers to entry?

\begin{enumerate}
\item For HM erasing reputation hurts incumbent more (reputational advantage is more important than data advantage) (Why?)
\item For HMR/SM - erasing either hurts a bit, but erasing both hurts a lot implying that reputational and data advantage substitute for one another
\end{enumerate}

\section*{Story}

\begin{itemize}
\item Competition in learning environments - compete for users based on quality alone but quality of different actions is unknown and need to learn the quality of actions (modeled as an iid MAB). Learn by committing to a learning algorithm at the start of the world. Want the adoption of traditionally ``better" algorithms since it is socially beneficial but when does this happen?
\item We introduce a simple metric of reputation - agents want to maximize their reward and use a reputation score as a proxy for that. Reputation score is a sliding window average of the reward experienced by past agents that had selected this firm. Agents are myopic and non-strategic so they do not consider other factors such as the algorithm employed by the firms.
\item \textbf{Problem}: Competition imposes incentive compatibility constraints on myopic consumers (i.e. firm needs to choose actions so that they actually get selected by users in competition). Start with consumers using the HardMax decision rule.
\item Firms face a tradeoff between picking algorithms that maximize short-term reputation and those engage in pure exploration (which assists in maximizing long-term reputation). Consumers don't care about information gain from exploration but do care about reputation. Thus, algorithms that engage in pure exploration suffer short-term reputational costs and lose in the competition game to algorithms that maximize short-term reputation. Another aspect of this is that it depends on the hardness of the learning instance - the ``harder" learning instances have larger (short-term) reputational costs but potentially larger (long-run) reputational advantage.
\item How to incentivize firms better algorithms? Sufficient number of consumers whose incentive compatibility constraints are effectively removed. A few ways this can happen
\begin{enumerate}
\item Sufficient number of periods as the monopolist in the market (incumbent experiment results)
\item ``Random agents" or probabilistic decision rules (HMR or SM response)
\item ``Warm start" or free agents at the start of the game
\end{enumerate}
\item Focusing on the monopoly case and HM decision rule, the firm accumulates both a reputational advantage and a data advantage. Which is more important in the competition game for serving as barriers to entry. We see that it is reputation (I have no good intuition for this yet).
\item \textbf{Takeaway}: competition can stifle ``innovation" if learning has reputational effects due to incentive constraints on consumers. Counterintuitively, monopoly can have benefit for innovation since it allows for exploration by not having to worry about incentive constraint of consumers. However, sometimes even a traditionally ``better" algorithm according to mean regret cannot win in competition since relative reputation matters more.

\end{itemize}
\end{document}