### 2018-08-31 paper/content/sim_details.tex
Rewrote "simulation details".

\documentclass[../competing_bandits.tex]{subfiles}
\begin{document}

%\subsubsection{Simulation details.}
%\label{section:3}

\subsubsection{MAB instances.} 
In all experiments, there are $K=10$ arms. Since we focus on 0-1 rewards, an instance of the MAB problem is specified by the mean reward vector $(\mu(a):\; a\in A)$, henceforth \MRV. Initially the \MRV is drawn from some distribution, termed \emph{MAB instance}. We consider several representative MAB instances.
\begin{enumerate}
\item \emph{Needle In Haystack}: one arm (the ``needle") is chosen uniformly at random. This arm has mean reward $.7$, and the remaining ones have mean reward $.5$.

\item \emph{Uniform instance}: the mean reward of each arm is drawn independently and uniformly from $[0.25, 0.75]$.
\item \emph{Heavy-tail instance}: the mean reward of each arm is drawn independently from $\Beta(.6,.6)$ distribution.%
    \footnote{This distribution has substantial ``tail probabilities".}
\end{enumerate}

\subsubsection{Simulation details.} 
For each MAB instance we draw $N = 1000$ \MRVs independently from the corresponding distribution. We use this same collection of \MRVs for all experiments with this MAB instance. For each \MRV we draw a table of realized rewards (\emph{realization table}), and use this same table for all experiments on this \MRV. This ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings. More specifically, the realization table is a 0-1 matrix $W$ with $K$ columns which correspond to arms, and $T+T_{\max}$ rows, which correspond to rounds. Here $T_{\max}$ is the maximal duration of the ``warm start" in our experiments, \ie the maximal value of $X+T_0$. For each arm $a$, each value $W(\cdot,a)$ is drawn independently from Bernoulli distribution with expectation $\mu(a)$. Then in each experiment, the reward of this arm in round $t$ of the warm start is taken to be $W(t,a)$, and its reward in round $t$ of the game is $W(T_{\max}+t,a)$.

\OMIT{This realization table, as well as fixing the random seed for the same bandit instance and realization table across experiments, ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings. In the competition game we draw from the $T \times K$ portion of the table, so that if two different algorithms picks arm $a$ at time $t$, they get the same $[a, t]$ realization in the table. This setup also ensures that in the warm start period, increasing the warm start from $k$ to $k + 10$ results in the same behavior in the first $k$ rounds.}

We fix the sliding window size $M = 100$. We found that low values of $M$ induced too much random noise in the results, and increasing $M$ further did not make a substantial qualitative difference.

While we experiments with various MAB instances and parameter settings, in the main paper we only report plots for selected, representative experiments. Additional plots can be found in the supplement. Unless noted otherwise, our findings are based on and consistent with all these plots.

\end{document}