### 2018-09-05 paper/content/barriers.tex
more table placements

\documentclass[../competing_bandits.tex]{subfiles}
\begin{document}

\section{Data and Reputation as Barriers to Entry}\label{section:6}

In the temporary monopoly setting, the incumbent firm enjoys a first
mover advantage, which allows the firm to acquire data on potentially
suboptimal actions without incurring an immediate reputational
consequence, and to build up a high reputation before the entrant
enters the market. Thus, the early entry gives the incumbent both a
\textit{data} advantage and a \textit{reputational} advantage over the
entrant. This provides a quantitative perspective on the classic
``first mover advantage" \cite{kerin1992first} in the digital economy
where data can serve as a barrier to entry in many markets.

While the two advantages together serve as strong barriers to entry, a
natural question to ask is whether either one of the advantages alone
suffices as a strong barrier to entry. If both of them are barriers of
entry, which of the two is the stronger?

To answer these questions, we isolate the effect from each of the two
advantages and run two additional experiments. In the first experiment
(called \emph{reputation erased experiment}), the reputation of the
incumbent is artificially reset upon the entry of the entrant {so that
  the incumbent's reputation score is only based on the rewards
  generated in the warm start period}. In the second experiment
(called \emph{information erased experiment}), the information gained
by the incumbent is erased upon the entry of the entrant in the sense
that its posterior is reset to the initial prior. The results of the
experiments are presented in Table~\ref{barrier_exp}, which lead to
our Finding~\ref{barrier-find}.

%\begin{table}[ht]
%\centering
%\begin{tabular}{cc|c|c|c|}
%  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
%  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
%            & $\TS$ & \makecell{\textbf{0.021} \\$\pm$0.009}& 		\makecell{\textbf{0.16} \\$\pm$0.02} &
%            \makecell{\textbf{0.21} \\ $\pm$0.02} \\ \cline{3-5}
%Entrant  & $\DEG$ & \makecell{\textbf{0.26} \\$\pm$0.03} & \makecell{\textbf{0.3} \\$\pm$0.02} &
%\makecell{\textbf{0.26} \\ $\pm$0.02} \\ \cline{3-5}
%            & $\DG$ & \makecell{\textbf{0.34} \\ $\pm$0.03} & \makecell{\textbf{0.4} \\$\pm$0.03} &
%            \makecell{\textbf{0.33} \\$\pm$0.02} \\\cline{3-5}
%\end{tabular}
%\caption{Information Erased Experiment Heavy Tail}
%\label{info_erase}
%\end{table}
%
%\begin{table}[ht]
%\centering
%\begin{tabular}{cc|c|c|c|}
%  & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Incumbent} \\
%  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{$\TS$}  & \multicolumn{1}{c}{$\DEG$}  & \multicolumn{1}{c}{$\DG$} \\\cline{3-5}
%            & $\TS$ & \makecell{\textbf{0.0096} \\$\pm$0.006}& 		\makecell{\textbf{0.11} \\$\pm$0.02} &
%            \makecell{\textbf{0.18} \\$\pm$0.02} \\ \cline{3-5}
%Entrant  & $\DEG$ & \makecell{\textbf{0.073} \\ $\pm$0.01} & \makecell{\textbf{0.29} \\ $\pm$0.02} &
%\makecell{\textbf{0.25} \\$\pm$0.02} \\ \cline{3-5}
%            & $\DG$ &\makecell{\textbf{0.15} \\$\pm$0.02} & \makecell{\textbf{0.39} \\$\pm$0.03} &
%           \makecell{\textbf{0.33} \\ $\pm$0.02} \\\cline{3-5}
%\end{tabular}
%\caption{Reputation Erased Experiment Heavy Tail}
%\label{rep_erase}
%\end{table}

\begin{finding}\label{barrier-find}
\swedit{During the rounds in which the entrant is present, the incumbent
  with either data or reputational advantage alone gets a larger share
  than the share \gaedit{in the permanent duopoly, regardless of the algorithm she plays}.  When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two
  barriers.}
\OMIT{\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a larger market share compared to the share they get under permanent duopoly}.\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}}
\end{finding}

% \swcomment{for the first part of the finding, is it true for any algorithm played by the incumbent?}

\gaedit{While the data and reputation advantages seem to serve as substitutes for \DEG and \DG, it is interesting that under \TS the data advantage is larger. The reason for this is that, when playing \TS, not only does the early entry allow the incumbent to gather more data than the entrant, but it allows it to gather data on actions \textit{that may be hard to gather data on in competition}. In contrast, purposeful exploration in competition may have immediate adverse reputational consequences. However, in the incumbency period the incumbent faces no immediate reputational consequences from exploration and so has data on a more diversely distributed set of actions but has time to recover lost reputation from selecting the bad arms. As a result, the data advantage is larger for \TS precisely because the advantage that data gives is not just about the quantity of data but also about its distribution across actions.

Why is it that data and reputation alone provide such large barriers to entry? Having more data allows the incumbent to select better actions which allows her to quickly get a high reputation compared to an entrant that both needs to learn and work on building up a reputation. Starting with a large reputation score gives the incumbent a cushion for doing exploration and learning without facing as stark reputational consequences. It is interesting that, for \DEG and \DG, these two effects seem to serve nearly as perfectly substitutes.
}

\gadelete{
The reason \gadelete{for the finding} \gaedit{that we see that the data advantage is greater
under $\TS$} is that the early entry allows the incumbent to do
exploration without immediate reputational consequence in the early
rounds and so, even when reputation is erased, the incumbent has
already acquired sufficient information that simply exploiting that
information allows her to quickly recover her
reputation. \swcomment{which part are you explaining?} However, while
the reputational advantage gives the incumbent some cushion for doing
exploration the incumbent is still vulnerable to getting unlucky while
doing exploration and suffering the reputational consequences. Under all of the incumbent algorithms,
the incumbent obtains a data advantage by the mere fact that it is around in the market and gets data from consumers. However, we see that the data advantage is stronger under \TS since \TS acquires data that is hard to acquire under competition.
We only observe this under $\TS$ since
\TS not only acquiresThis
only happens under $\TS$ since $\TS$, by construction, is more
intelligent about the data that it acquires compared to $\DEG$ and
$\DG$.\swcomment{more intelligent about the data means?}
}

\OMIT{Though the setup is purely experimental, it is nonetheless interesting to look at if the same strategies remain best responses in this setting compared to the setting where both data and reputation are retained. We find that the best responses remain the same for the Heavy Tail instance, but for the Uniform instance they change where $\TS$ is weakly dominant for the incumbent under the reputation erased treatment but not under the information erased treatment.}

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../competing_bandits"
%%% End: