### 2018-09-05 paper/content/perf_in_iso.tex
new figure + clean up some edits

\documentclass[../competing_bandits.tex]{subfiles}
\begin{document}

\section{Algorithms' Performance in Isolation}\label{sec:isolation}

We start with a pilot experiment in which we investigate each algorithm's performance ``in isolation": in a stand-alone MAB problem without competition. We focus on reputation scores generated by each algorithm. We confirm that algorithms' performance is ordered as we'd expect:
    $\TS > \DEG > \DG$
for a sufficiently long time horizon.
For each algorithm and each MAB instance, we compute the mean reputation score at each round, averaged over all \MRVs. We plot the \emph{mean reputation trajectory}: how this score evolves over time. Figure \ref{prelim_means} shows such a plot for the Needle In Haystack instance; for other MAB instances the plots are similar. We summarize this finding as follows:

\begin{finding}
\textit{The mean reputation trajectories are arranged as predicted by prior work:
    $\TS > \DEG > \DG$ for a sufficiently long time horizon.}
\end{finding}

\begin{figure}
\includegraphics[scale=0.35]{figures/nih_iso_mean}
%\caption*{\tiny{Mean reputation trajectory The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs. The shaded area display 95\% confidence intervals.}}
\caption{Mean reputation trajectories for Needle-in-Haystack. The shaded area displays 95\% confidence intervals.}
\label{prelim_means}
\end{figure}

The mean reputation trajectory is probably the most natural way to represent an algorithm's performance on a given MAB instance. However, we found that the outcomes of the competition game are better explained with a different ``performance-in-isolation" statistic that is more directly connected to the game. Consider the  performance of two algorithms, Alg1 and Alg2, ``in isolation" on a particular MAB instance. The \emph{relative reputation} of Alg1 (vs. Alg2) at a given time $t$ is the fraction of \MRVs /realization tables for which Alg1 has a higher reputation score than Alg2. The intuition is that agent's selection in our model depends only on the comparison between the reputation scores.

\begin{figure}[ht]
\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
\includegraphics[scale=0.35]{figures/relative_nih_ts_dg.png}
%\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $\TS \geq \DG$. The shaded area display 95\% confidence intervals.}}
\caption{Relative reputation trajectory for $\TS$ vs $\DG$, on Uniform instance (top) and Needle-in-Haystack instance (bottom). Shaded area display 95\% confidence intervals.}
\label{relative_rep_plots}
\end{figure}

This angle allows a more nuanced analysis of reputation costs vs. benefits under competition. Figure \ref{relative_rep_plots} (top) shows the relative reputation trajectory for $\TS$ vs $\DG$ for the Uniform instance. The relative reputation is less than $\tfrac12$ in the early rounds, meaning that $\DG$ has a higher reputation score in a majority of the simulations, and more than $\tfrac12$ later on. The reason is the exploration in \TS leads to worse decisions initially, but allows for better decisions later. The time period when relative reputation vs. \DG dips below $\tfrac12$ can be seen as an explanation for the competitive disadvantage of exploration. Such period also exists for the Heavy-tail MAB instance. However, it does not exist for the Needle-in-Haystack instance, see Figure \ref{relative_rep_plots} (bottom).%
\footnote{We see two explanations for this: $\TS$ identifies the best arm faster  for the Uniform instance, and there are no ``very bad" arms which make exploration very expensive in the short term.}

\begin{finding}\label{find:period}
\textit{Exploration can lead to relative reputation vs. $\DG$ going below $\tfrac12$ for some initial time period. This happens for some MAB instances but not for some others.}
\end{finding}

\begin{definition}
For a particular MAB algorithm, a time period when relative reputation vs. \DG goes below $\tfrac12$ is called {\em exploration disadvantage period}. An MAB instance is called \emph{exploration-disadvantaged} if such period exists.
\end{definition}

Thus, Uniform and Heavy-tail instance are exploration-disadvantaged, but Needle-in-Haystack is not.

\end{document}