### 2019-06-17 journal paper/itcs18 paper/sec-related-work.tex

Multi-armed bandits (\emph{MAB}) is a particularly elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information. MAB problems have been studied in Economics, Operations Research and Computer Science for many decades; see \citep{Bubeck-survey12,Gittins-book11,slivkins-MABbook} for background on regret-minimizing and Bayesian formulations, respectively. A discussion of industrial applications of MAB can be found in \citet{MWT-WhitePaper-2016}.

The literature on MAB is vast and multi-threaded. The most related
thread concerns regret-minimizing MAB formulations with IID rewards
\citep{Lai-Robbins-85,bandits-ucb1}. This thread includes ``smart" MAB
algorithms that combine exploration and exploitation, such as UCB1
\citep{bandits-ucb1} and Successive Elimination
\citep{EvenDar-icml06}, and ``naive'' MAB algorithms that separate
exploration and exploitation, including explore-first and
$\eps$-Greedy \citep[\eg see][]{slivkins-MABbook}.

The three-way tradeoff between exploration, exploitation and incentives has been studied in several other settings:
incentivizing exploration in a recommendation system
    \citep{Che-13,Frazier-ec14,Kremer-JPE14,ICexploration-ec15,Bimpikis-exploration-ms17,Bahar-ec16,ICexplorationGames-ec16-working},
dynamic auctions
    \cite[\eg][]{AtheySegal-econometrica13,DynPivot-econometrica10,Kakade-pivot-or13},
pay-per-click ad auctions with unknown click probabilities
    \cite[\eg][]{MechMAB-ec09,DevanurK09,Transform-ec10-jacm},
coordinating search and matching by self-interested agents
    \citep{Bobby-Glen-ec16},
as well as human computation
    \cite[\eg][]{RepeatedPA-ec14,Ghosh-itcs13,Krause-www13}.

\citet{Bolton-econometrica99,Keller-econometrica05,Johari-ec12} studied models with self-interested agents jointly performing exploration, with no principal to coordinate them.

There is a superficial similarity (in name only) between this paper and the line of work on ``dueling bandits"
    \citep[\eg][]{Yue-dueling12,Yue-dueling-icml09}.
The latter is not about competing bandit algorithms, but rather about scenarios where in each round two arms are chosen to be presented to a user, and the algorithm only observes which arm has ``won the duel".

Our setting is closely related to the ``dueling algorithms" framework \citep{DuelingAlgs-stoc11} which studies competition between two principals, each running an algorithm for the same problem. However, this work considers algorithms for offline / full input scenarios, whereas we focus on online machine learning and the explore-exploit-incentives tradeoff therein. Also, this work specifically assumes binary payoffs (\ie win or lose) for the principals.

\xhdr{Other related work in economics.} The competition vs. innovation relationship and the inverted-U shape thereof have been introduced in a classic book \citep{Schumpeter-42}, and remained an important theme in the literature ever since \cite[\eg][]{Aghion-QJE05,Vives-08}. Production costs aside, this literature treats innovation as a priori beneficial for the firm. Our setting is very different, as innovation in exploration algorithms may potentially hurt the firm.

A line of work on \emph{platform competition}, starting with \cite{Rysman09}, concerns competition between firms (\emph{platforms}) that improve as they attract more users (\emph{network effect}); see \citet{Weyl-White-14} for a recent survey. This literature is not concerned with \innovation, and typically models network effects exogenously, whereas in our model network effects are endogenous: they are created by MAB algorithms, an essential part of the model.

Relaxed versions of rationality similar to ours are found in several notable lines of work. For example, ``random agents" (a.k.a. noise traders) can side-step the ``no-trade theorem'' \citep{Milgrom-Stokey-82}, a famous impossibility result in financial economics. The \SoftMaxRandom model is closely related to the literature on \emph{product differentiation}, starting from \cite{Hotelling-29}, see \cite{Perloff-Salop-85} for a notable later paper.

There is a large literature on non-existence of equilibria due to small deviations   (which is related to the corresponding result for \HardMaxRandom), starting with \cite{Rothschild-Stiglitz-76} in the context of health insurance markets. Notable recent papers \citep{Veiga-Weyl-16,Azevedo-Gottlieb-17} emphasize the distinction between \HardMax and versions of \SoftMaxRandom.

% moved this thought to the intro.
\OMIT{While agents' rationality and severity of competition are often modeled separately in the literature, it is not unusual to have them modeled with the same ``knob" \cite[\eg][]{Gabaix-16}.}

%%% Local Variables:
%%% TeX-master: "main.tex"
%%% End: