### 2019-08-14 journal paper/itcs18paper/sec-model.tex

\xhdr{Principals and agents.} There are two principals and $T$ agents. The game proceeds in rounds (we will sometimes refer to them as \emph{global rounds}). In each round $t\in [T]$, the following  interaction takes place. A new agent arrives and chooses one of the two principals. The principal chooses a recommendation: an action $a_t\in A$, where $A$ is a fixed set of actions (same for both principals and all rounds). The agent follows this recommendation, receives a reward $r_t\in [0,1]$, and reports it back to the principal.

The rewards are i.i.d. with a common prior. More formally, for each action $a\in A$ there is a parametric family $\psi_a(\cdot)$ of
reward distributions, parameterized by the mean reward $\mu_a$. (The paradigmatic case is 0-1 rewards with a given expectation.) The
mean reward vector $\mu = (\mu_a:\; a\in A)$ is drawn from prior distribution $\priorMu$ before round $1$. Whenever a given action $a\in A$ is chosen, the reward is drawn independently from distribution $\psi_a(\mu_a)$. The prior $\priorMu$ and the distributions $(\psi_a(\cdot)\colon a\in A)$ constitute the (full) Bayesian prior on rewards, denoted $\prior$.

Each principal commits to a learning algorithm for making recommendations. This algorithm follows a protocol of \emph{multi-armed bandits} (\emph{MAB}). Namely, the algorithm proceeds in time-steps:%
\footnote{These time-steps will sometimes be referred to as \emph{local steps/rounds}, so as to distinguish them from ``global rounds" defined before. We will omit the local vs. local distinction when clear from the context.} each time it is called, it outputs a chosen action $a\in A$ and then inputs the reward for this action. The algorithm is called only in global rounds when the corresponding principal is chosen.

The information structure is as follows. The prior $\prior$ is known to everyone. The mean rewards $\mu_a$ are not revealed to anybody. Each agent knows both principals' algorithms, and the global round when (s)he arrives, \emph{but not} the rewards of the previous agents. Each principal is completely unaware of the rounds when the other is chosen.

\xhdr{Some terminology.} The two principals are called ``Principal
1" and ``Principal 2".
%Principal $i\in \{1,2\}$ is also denoted \prin[i].
The algorithm of principal $i\in \{1,2\}$ is called ``algorithm $i$" and denoted
\alg[i]. The agent in global round $t$ is called ``agent $t$"; the
chosen principal is denoted $i_t$.

Throughout, $\E[\cdot]$ denotes expectation over all applicable randomness.

\xhdr{Bayesian-expected rewards.}
Consider the performance of a given algorithm \alg[i], $i\in \{1,2\}$, when it is run in isolation (\ie without competition, just as a bandit algorithm). Let $\rew_i(n)$ denote its Bayesian-expected reward for the $n$-th step.

Now, going back to our game, fix global round $t$ and let $n_i(t)$ denote the number of global rounds before $t$ in which this principal is chosen. Then:
\begin{align*}
 \E[r_t \mid \text{principal $i$ is chosen in round $t$ and $n_i(t)=n$} ]
    = \rew_i(n+1) \quad (\forall n\in\N).
\end{align*}

\xhdr{Agents' response.}
Each agent $t$ chooses principal $i_t$ as as follows: it chooses a distribution over the principals, and then draws independently from this distribution. Let $p_t$ be the probability of choosing principal $1$ according to this distribution. Below we specify $p_t$; we need to be careful so as to avoid a circular definition.

Let $\mI_t$ be the information available to agent $t$ before the
round. Assume $\mI_t$ suffices to form posteriors for quantities
$n_i(t)$, $i\in \{1,2\}$, denote them by $\posteriorN{i}{t}$. Note
that the Bayesian expected reward of each principal $i$ is a function
only of the number rounds he was chosen by the agents, so the
posterior mean reward for each principal $i$ can be written as
\begin{align*}
 \PMR_i(t) := \E[r_t \mid \text{$\mI_t$ and $i_t=i$} ]
    = \E[\rew_i(n_i(t)+1) \mid \mI_t]
    = \E_{n\sim \posteriorN{i}{t}}[\rew_i(n+1)].
\end{align*}
This quantity represents the posterior mean reward for principal $i$
at round $t$, according to information $\mI_t$; hence the notation \PMR. In general, probability $p_t$ is defined by the
posterior mean rewards $\PMR_i(t)$ for both principals. We assume a
somewhat more specific shape:
\begin{align}
p_t = \respF\left(\; \PMR_1(t) - \PMR_2(t) \;\right).
%p_t = \respF(\Delta_t), \quad \text{where } \Delta_t := \PMR_1(t) - \PMR_2(t).
\end{align}
Here $\respF:[-1,1]\to [0,1]$ is the \emph{response function}, which is the same for all agents. We assume that the response function is known to all agents.

To make the model well-defined, it remains to argue that information $\mI_t$ is indeed sufficient to form posteriors on $n_1(t)$ and $n_2(t)$. This can be easily seen using induction on $t$.

Since all agents arrive with identical information (other than knowing which global round they arrive in), it follows that all agents have identical posteriors for $n_{i,t}$ (for a given principal $i$ and a given global round $t$). This posterior is denoted $\posteriorN{i}{t}$.

\xhdr{Response functions.}
We use the response function $\respF$ to characterize the amount of rationality and competitiveness in our model. We assume that $\respF$ is monotonically non-decreasing, is larger than $\tfrac12$ on the interval $(0,1]$, and smaller than $\tfrac12$ on the interval $[-1,0)$. Beyond that, we consider three specific models, listed in the order of decreasing rationality and competitiveness (see \reffig{fig:response-functions}):

\begin{figure}
\begin{center}
  \begin{tikzpicture}[scale=2.0]
    \draw[->] (-1.1,0) -- (1.1,0) node[right] {$\Delta_t := \PMR_1(t) - \PMR_2(t)$};
    \draw[->] (0,-0.1) -- (0,1.1) node[above]
        {$p_t = \text{prob. of choosing principal 1}$};
    \draw[scale=1.0,domain=-1:0,smooth,variable=\q,blue, line width=0.50mm] plot ({\q},{0});
    \draw[scale=1.0,domain=0:1,smooth,variable=\q,blue,line width=0.50mm] plot ({\q},{1});
    \draw[scale=1.0,domain=-1:0,smooth,variable=\y,red]  plot ({\y},{0.1});
    \draw[scale=1.0,domain=0:1,smooth,variable=\y,red]  plot ({\y},{0.9});
    \draw[scale=1.0,domain=-1:1,smooth,variable=\y,cyan, line width=0.45mm, dash pattern=on 3pt off 2pt]  plot ({\y},{1/(1 + 1/(9^\y))});
    % \node[above, blue] at (0.5, 0.5) {\footnotesize $2 q (1 - q)$};
     \node[left] at (0, 0.5) {\footnotesize $1/2$};
     \node[left] at (0, 1) {\footnotesize $1$};
     \node[below left] at (0, 0) {\footnotesize $0$};
     \node[below ] at (1, 0) {\footnotesize $1$};
     \node[below ] at (-1, 0) {\footnotesize $-1$};
  \end{tikzpicture}
\end{center}
\caption{The three models for agents' response function: \HardMax is thick blue, \HardMaxRandom is slim red, and \SoftMaxRandom is the dashed curve.}
\label{fig:response-functions}
\end{figure}

\begin{OneLiners}
\item \HardMax: $\respF$ equals $0$ on the interval $[-1,0)$ and $1$
  on the interval $(0,1]$. In other words, the agents will
  deterministically choose the principal with the higher posterior
  mean reward.

\item \HardMaxRandom:
    % $\respF$ equals $\eps$ on the interval $[-1,0)$ and $1-\eps'$ on the interval $(0,1]$, where $\eps,\eps'\in (0,\tfrac12)$ are some positive constants. In words, each agent is a \HardMax agent with probability $1-\eps-\eps'$, and with the remaining probability she makes a random choice.
    $\respF$ equals $\eps_0$ on the interval $[-1,0)$ and $1-\eps_0$ on the interval $(0,1]$, where $\eps_0\in (0,\tfrac12)$ are some positive constants. In words, each agent is a \HardMax agent with probability $1-2\eps_0$, and with the remaining probability she makes a random choice.

\item \SoftMaxRandom: $\respF(\cdot)$ lies in the interval $[\eps_0,1-\eps_0]$, $\eps_0>0$, and is ``smooth" around $0$ (in the sense defined precisely in Section~\ref{sec:soft}).
\end{OneLiners}
% \asedit{(We assume $\respF(-1)+\respF(1)=1$ in \HardMaxRandom and \SoftMaxRandom only to simplify notation.)}

We say that $\respF$ is \emph{symmetric} if $\respF(-x)+\respF(x)=1$
for any $x\in [0,1]$. This implies \emph{fair tie-breaking}:
$\respF(0)=\tfrac12$.% , and $\eps=\eps'$ in the
% definitions above.

\xhdr{MAB algorithms.}
We characterize the inherent quality of an MAB algorithm in terms of its \emph{Bayesian Instantaneous Regret} (henceforth, \BIR), a standard notion from machine learning:
\begin{align}
\BIR(n) := \E_{\mu\sim\priorMu}
    \left[ \max_{a\in A} \mu_a\right] - \rew(n),
\end{align}
where $\rew(n)$ is the Bayesian-expected reward of the algorithm for the $n$-th step, when the algorithm is run in isolation. We are primarily interested in how \BIR scales with $n$; we treat $K$, the number of arms, as a constant unless specified otherwise.

We will emphasize several specific algorithms or classes thereof:

\begin{itemize}
\item ``smart" MAB algorithms that combine exploration and exploitation, such as UCB1 \cite{bandits-ucb1} and Successive Elimination \cite{EvenDar-icml06}. These algorithms achieve
        $\BIR(n) \leq  \tilde{O}(n^{-1/2})$
    for all priors and all (or all but a very few) steps $n$. This bound is known to be tight for any fixed $n$.
    \footnote{This follows from the lower-bound analysis in \cite{bandits-exp3}.}

\item ``naive" MAB algorithms that separate exploration and exploitation, such as Explore-then-Exploit and $\eps$-Greedy. These algorithms have dedicated rounds in which they explore by choosing an action uniformly at random. When these rounds are known in advance, the algorithm suffers constant \BIR in such rounds. When the ``exploration rounds" are instead randomly chosen by the algorithm, one can usually guarantee an inverse-polynomial upper bound \BIR, but not as good as the one above: namely,
        $ \BIR(n) \leq \tilde{O}(n^{-1/3})$.
    This is the best possible upper bound on \BIR for the two algorithms mentioned above.

\item \DynGreedy: at each step, recommends the best action according to the current posterior: an action $a$ with the highest posterior expected reward
        $\E[\mu_a \mid \mI]$,
     where $\mI$ is the information available to the algorithm so far.
     \DynGreedy has (at least) a constant \BIR for some reasonable priors, \ie
        $\BIR(n)>\Omega(1)$.

\item \StaticGreedy: always recommends the prior best action,\ie an action $a$ with the highest prior mean reward  $\E_{\mu\sim \priorMu}[\mu_a]$. This algorithm typically has constant \BIR.
\end{itemize}

We focus on MAB algorithms such that $\BIR(n)$ is non-increasing; we call such algorithms \emph{monotone}. While some reasonable MAB algorithms may occasionally violate monotonicity, they can usually be easily modified so that monotonicity violations either vanish altogether, or only occur at very specific rounds (so that agents are extremely unlikely to exploit them in practice).

More background and examples can be found in Appendix~\ref{app:examples}. In particular, we prove that \DynGreedy is monotone.

\xhdr{Competition game between principals.}
Some of our results explicitly study the game between the two principals. We model it as a simultaneous-move game: before the first agent arrives, each principal commits to an MAB algorithm. Thus, choosing a pure strategy in this game corresponds to choosing an MAB algorithm (and, implicitly, announcing this algorithm to the agents).

Principal's utility is primarily defined as the market share, \ie the number of agents that chose this principal. Principals are risk-neutral, in the sense that they optimize their expected utility.

\xhdr{Assumptions on the prior.} We make some technical assumptions for the sake of simplicity. First, each action $a$ has a positive probability of being the best action according to the prior:
\begin{align}\label{eq:assn-prob}
\forall a\in A:\;\;\;  \Pr_{\mu\sim \priorMu}[\mu_a  > \mu_{a'}
\;\forall a'\in A]
> 0.
\end{align}

Second, posterior mean rewards of actions are pairwise distinct almost
surely. That is, the history $h$ at any step of an MAB algorithm%
\footnote{The \emph{history} of an MAB algorithm at a given step
  comprises the chosen actions and the observed rewards in all
  previous steps in the execution of this algorithm.}
 satisfies
\begin{align}\label{eq:assn-distinct}
    \E[\mu_a \mid h] \neq \E[\mu_{a'}\mid h] \quad \forall a,a'\in A,
\end{align}
\asedit{except at a set of histories of probability $0$}.  In
particular, prior mean rewards of actions are pairwise distinct:
$\E[\mu_a] \neq \E[\mu_a']$ for any $a,a'\in A$.

We provide two examples for which property \eqref{eq:assn-distinct} is `generic', in the sense that it can be enforced almost surely by a
  small random perturbation of the prior. Both examples focus on 0-1 rewards and priors $\priorMu$ that are independent across arms. The first example
  assumes Beta priors on the mean rewards, and is very easy.%
  \footnote{Suppose the rewards are Bernouli r.v. and the mean reward
    $\mu_a$ for each arm $a$ is drawn from some Beta distribution
    $\text{Beta}(\alpha_a, \beta_a)$. Given any history that contains
    $h_a$ number of heads and $t_a$ number of tails from arm $a$, the
    posterior mean reward is
    $\frac{\alpha_a + h_a}{\alpha_a + h_a + \beta_a + t_a}$. Note that
    $h_a$ and $t_a$ take integer values. Therefore, perturbing the
    parameters $\alpha_a$ and $\beta_a$ independently with any
    continuous noise will induce a prior with property
    \eqref{eq:assn-distinct} with probability 1.  }  The second
  example assumes that mean rewards have a
  finite support, see Appendix~\ref{app:perturb} for details.

\xhdr{Some more notation.} Without loss of generality, we label actions as $A=[K]$ and sort them according to their prior mean rewards, so that
    $ \E[\mu_1] > \E[\mu_2] > \ldots > \E[\mu_K]$.

Fix principal $i\in \{1,2\}$ and (local) step $n$. The arm chosen by algorithm \alg[i] at this step is denoted $a_{i,n}$, and the corresponding \BIR is denoted $\BIR_i(n)$. History of \alg[i] up to this step is denoted $H_{i,n}$.
%The posterior distribution of agent $t$ for quantity $n_t(t)$ is denoted $\posteriorN{i}{t}$.

Write
    $\PMR(a\mid E) = \E[\mu_a \mid E]$
for posterior mean reward of action $a$ given event $E$.

%\end{assumption}

\subsection{Generalizations}
\label{sec:model-extensions}

Our results can be extended compared to the basic model described above.

%First, the two principals may have different action sets, as long as
%    $\E\left[ \max_{a\in A} \mu_a\right]$,
%the Bayesian-expected reward of the best action, is the same for both principals.

First, unless specified otherwise, our results allow a more general notion of principal's utility that can depend on both the market share and agents' rewards. Namely, principal $i$ collects $U_i(r_t)$ units of utility in each global round $t$ when she is chosen (and $0$ otherwise), where $U_i(\cdot)$ is some fixed non-decreasing function with $U_i(0)>0$. In a formula,
\begin{align}\label{eq:general-utility}
\textstyle U_i := \sum_{t=1}^T\; \indicator{i_t=i}\cdot U_i(r_r).
\end{align}

Second, our results carry over, with little or no modification of the proofs, to much more general versions of MAB, as long as it satisfies the i.i.d. property. In each round, an algorithm can see a \emph{context} before choosing an action (as in \emph{contextual bandits}) and/or additional feedback other than the reward after the reward is chosen (as in, \eg \emph{semi-bandits}), as long as the contexts are drawn from a fixed distribution, and the (reward, feedback) pair is drawn from a fixed distribution that depends only on the context and the chosen action. The Bayesian prior $\prior$ needs to be a more complicated object, to make sure that \PMR and \BIR are well-defined. Mean rewards may also have a known structure, such as Lipschitzness, convexity, or linearity; such structure can be incorporated via $\prior$. All these extensions have been studied extensively in the literature on MAB, and account for a substantial segment thereof; see \cite{Bubeck-survey12} for background and details.

\subsection{Chernoff Bounds}

We use an elementary concentration inequality known as {\em Chernoff Bounds}, in a formulation from~\cite{MitzUpfal-book05}.
\begin{theorem}[Chernoff Bounds]
\label{thm:chernoff}
Consider $n$ i.i.d. random variables $X_1 \ldots X_n$ with values in $[0,1]$. Let
    $X = \tfrac{1}{n} \sum_{i=1}^n X_i$ be their average, and let $\nu = \E[X]$. Then:
\[ \min\left(\; \Pr[ X-\nu > \delta \nu ],\quad
                \Pr[ \nu-X > \delta \nu ]
    \; \right)
    < e^{-\nu n \delta^2/3}
    \quad \text{for any $\delta\in (0,1)$.}
\]
\end{theorem}

%, and the choice of algorithm are commonly known
%to the agents. Given any finite class of MAB learning algorithms
%$\Pi$, a pure strategy for the principal $j$ in the competition game
%is an algorithm $\pi_j$ in the class $\Pi$ and a mixed strategy
%$\sigma_j$ is a probability distribution over $\Pi$. We will write
%$u_j(\sigma)$ and $u(\pi)$ to denote the expected utility for the
%principal $j$ in mixed strategy profile $\sigma$ and pure strategy
%profile $\pi$.  The principal gets a unit reward for being selected by
%an agent, so the utility function can be written as
%\[
%u_j(\pi) = \Ex{}{\sum_{t=1}^T\mathbf{1}[P^t = j]}
%\]
%A variant of the utility function we have is the number of agents
%weighted by the rewards for the agents
%\[
%u_j(\pi) = \Ex{}{\sum_{t=1}^T \mathbf{1}[P^t = j] r^t}
%\]
%\iffalse
%We consider a few models for the utility of the principals.  For
%example, the principal gets the reward realized by the agents it
%advised.
%\fi
%
%We will analyze the equilibrium strategies of the game under different
%environments.
%
%\begin{definition}[Equilibrium in the competition game]
%  A pure strategy profile $\pi$ is a pure strategy Nash equilibrium (PSNE) if
%  for any principal $j$ in the game and any algorithm $a\in \Pi$
%\[
%  u_j(\pi) \geq u_j(a, \pi_{-j}).
%\]
%A mixed strategy profile $\sigma$ is a (mixed strategy) Nash
%equilibrium if for any principal $j$ and any algorithm $a\in\Pi$,
%\[
%  \Ex{\sigma}{u_j(\sigma)} \geq u_j(a, \sigma_{-j}).
%\]
%\end{definition}
%
%We will measure the quality of an equilibrium by studying the welfare
%of the agents with respect to the system regret.

%%% Local Variables:
%%% TeX-master: "main.tex"
%%% End: